---
title: "Boot Strap the Covid"
author: "brooke"
date: "7/28/2022"
output: html_document
---

 **This document:** \
- creates a new data set from the covid and census data \
- conducts a hypothesis test \
- boot straps the new data set \
- visualizes the results


## Part One: create and examine a sample data set to bootstrap 
 
#### Import the Census and COVID-19 case data set created from the [Create Retrieve, Wrangle, Mutate, Merge](https://rpubs.com/ekoorb03/Guides_Create_Covid_Census) document, and select the metrics: unemployment_rate, covid_mortality_quintile
```{r sample data}
library(tidyverse)

boot_covid <- read_csv("~/STA 518/BrookemWalters-Portfolio/Stats 518 Final Project/Bootstrapping/covid_census_bs.csv") %>% 
  select(County, unemployment_rate, covid_mortality_quintile, Deaths_Per_Pop_Thousand) %>% 
  tibble()
```
 

### visualize the observations
Michigan County Unemployment Rate and COVID-19 deaths per 1000, Q1 = the lowest percentile & Q5 = the highest percentile for COVID-19 mortality 
It appears that mean unemployment rate rises with COVID-19 mortality rate. 
```{r box plot}
boxplot(boot_covid$unemployment_rate~boot_covid$covid_mortality_quintile, las = 1, ylab = "Unemployment Rate",
        xlab = "COVID-19 Mortality Quintile", main = "Unemployment Rate by COVID-19 Mortality Quintile")

```

 


### To keep the bootstrap simple, I'll just look at Q1 and Q5:
```{r filter the data}
boot_covid <- boot_covid %>% 
  filter(covid_mortality_quintile == 1 |covid_mortality_quintile == 5)
```



### Compare the two quintiles
```{r gg box plot}
boxplot(boot_covid$unemployment_rate~boot_covid$covid_mortality_quintile, las = 1, ylab = "Unemployment Rate", 
        xlab = "COVID-19 Mortality Quintile", main = "Unemployment Rate by COVID-19 Mortality Quintile")

```

## Hypothesis Testing
The code below conducts "Welch Two Sample t-test", I can compare the results to the bootstrap

**Specify the Null Hypothesis:** There is no difference in mean unemployment rate between Q1 and Q5 \
**Specify the Alternative Hypothesis:**  There is a difference in mean unemployment rate between Q1 and Q5 \
**Calculate the Test Statistic:**  T = -2.6822 \
**Calculate the P-Value:** p = 0.019 \
**Drawing a Conclusion:** Where alpha = 0.05, p < alpha,  reject H0, there is a difference in average unemployment rate based on COVID-19 mortality rate (Q1 vs Q5)

** However, my sample set **does not meet** the conditions for a two-sample T test, I would not use this as evidence, so I'll do a bootstrap and compare! 
```{r Welch Two Sample t-test}

t.test(boot_covid$unemployment_rate~boot_covid$covid_mortality_quintile, paired = FALSE, var.eq = FALSE)

```

## Part Two

### Calculate the Test Statistic for Boot Strapping  
```{r difference of unemployment means betwen the lowest and highest percentiles}
test_stat_one <-  abs(mean(boot_covid$unemployment_rate[boot_covid$covid_mortality_quintile == 5])) - mean(boot_covid$unemployment_rate[boot_covid$covid_mortality_quintile == 1])

round(test_stat_one, 2)
```

### Set a seed for reproducibility
```{r set seed}
set.seed(11062)
```



### Set up the bootstrap 
```{r setting up the bootstrap}
#n = number of observations
n <- length(boot_covid$unemployment_rate)

#b = number of bootstrap samples
B <- 10000

# assign the variable I'm testing a name for easier coding
variable <- boot_covid$unemployment_rate
```


### Collect the bootstrap samples: 33 by 10,000 matrix where each column is a bootstrap re-sample
```{r make the samples}
bootstrap_samples <-  matrix(sample(variable, size = n*B, replace = TRUE),
                             nrow = n, ncol = B)
```


```{r create vectors}
boot_test_stat_one <- rep(0,B)
```


###  Boot Strap Loop!
```{r calculate the boot-test-stat1 in a loop}
for (i in 1:B) {
boot_test_stat_one[i] <- abs(mean(bootstrap_samples [1:16,i]) - 
                               mean(bootstrap_samples [17:33,i]))
}

```

```{r looking at the first 20 bootstrap samples}
round(boot_test_stat_one[1:20],2)
```
### Taking a look at the bootstraps samples
```{r testing}
(boot_test_stat_one >= test_stat_one)[1:20]
```


time to calculate the bootstrap p-value
0 = False, 1 = True
```{r means of the true or false will yield the p-value}

mean(boot_test_stat_one >= test_stat_one)

```
**Specify the Null Hypothesis:** There is no difference in mean unemployment rate between Q1 and Q5 \
**Specify the Alternative Hypothesis:**  There is a difference in mean unemployment rate between Q1 and Q5 \
**Calculate the Test Statistic:**  T = 1.36 \
**Calculate the P-Value:** p = 0.0125 \
**Drawing a Conclusion:** Where alpha = 0.05, p < alpha,  reject H0, there is a difference in average unemployment rate based on COVID-19 mortality rate (Q1 vs Q5)

The bootstrap results yielded the same conclusion as the two-sample T test. 




###### ---[Back to Table of Contents](https://github.com/BrookemWalters/BrookemWalters-Portfolio#table-of-contents-brooke-walters-portfolio)---

