---
title: "Exploring the Data with Scatter Plots"
author: "brooke"
date: "7/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Exploring the Data With Scatter Plots"

I created four dynamic scatter plot tables that show the county by county relationships between  covid-19 morbity and :
- public housing assistance rate 
- median age
- median income
- higher education attainment
- 

the data will be exported into a CVS file and transformed into an aesthetically pleasing dynamic table 

[Here's](link) the final table!


The script below:

1) Divides the data into quintiles by covid-19 morbidity rate per 1,000 population
2) Updates variables names into a descriptive format
3) Create 4 interactive scatter plots of Michigan counties


## Step 0 
- To use plotly graphs on the GitHub page, you must sign up for plotly and issue an api key.
import plotly.express as px
import chart_studio
set your api
chart_studio.tools.set_credentials_file(username=username, api_key=api_key)


https://stackoverflow.com/questions/72378397/about-plotly-how-can-i-post-it-for-my-github-pages



## Step 1
 - Load plotly
#install package if needed

```{r load ploty}
library(plotly)
```


## Step 2
- create quintiles for an added layer the scatter plots

```{r create quintiles}
covid_quintile <- covid_census  %>% 
  mutate(quintile_group = ntile(Deaths_Per_Pop_Thousand,5))

```

1 = lowest Covid deaths per 1,000
5 = highest Covid deaths per 1,000


## Step 3
- Convert the quintile numbers into factors, so the variable can have descrete formatting in the plots


```{r convert to factors}
covid_quintile$quintile_group <-  covid_quintile$quintile_group %>%  as.factor 
```


## Step 4
- Convert data into tibble for ease of wrangling

```{r}
covid_quintile <- tibble(covid_quintile)
```


## Step 5
- Update variable names to be descriptive, the other set of descriptive names caused an error if used before this point due to the symbols used

```{r update names}
names(covid_quintile )[5] <- "Median Age"
names(covid_quintile )[6] <- "Median Household Income"
names(covid_quintile )[7] <- "% of Adults 25+ with College Degrees"
names(covid_quintile )[9] <- "% of Households on Public Assistance"
names(covid_quintile )[24] <- "Quintile"
names(covid_quintile )[23] <- "Covid Deaths Per Thousand"
names(covid_quintile )[22] <- "Total Covid Deaths"
```




## Steps 6 - 10
- create a scatter plot showing the  relationship between Covid Deaths per 1000, Total Covid Death, Quintiles, 
- update the x axis with the desired variable, and update the title to reflect the changes

THE LEGEND NEEDS TO BE FIXED TO REFLECT THE BUBBLE SIZE

```{r housing assistance plots}
assist_sp <- covid_quintile %>% 
  ggplot( aes(x = `% of Households on Public Assistance`, y = `Covid Deaths Per Thousand`, size = `Total Covid Deaths`, color=Quintile, text=County)) +
  geom_point(alpha = 6/10) +
  ggtitle("% of Michigan HHs on Assistance by Covid Deaths Per Thousand" ) +
 theme(panel.background = element_rect(fill = "white",
                                colour = "beige",
                                size = 0.5, linetype = "solid")) +
         
  scale_color_manual(values = c("#999999", "#00798c", "#66a182", "#edae49", "#dc0000b2")) 




ggplotly(assist_sp )
```
```{r median age plots}
age_sp <- covid_quintile %>% 
  ggplot( aes(x = `Median Age`, y = `Covid Deaths Per Thousand`, size = `Total Covid Deaths`, color=Quintile, text=County)) +
  geom_point(alpha = 6/10) +
  ggtitle("Median Age in Michigan Counties by Covid Deaths Per Thousand" ) +
 theme(panel.background = element_rect(fill = "white",
                                colour = "beige",
                                size = 0.5, linetype = "solid")) +
         
  scale_color_manual(values = c("#999999", "#00798c", "#66a182", "#edae49", "#dc0000b2")) 




ggplotly(age_sp )
```


```{r education plots}
ed_sp <- covid_quintile %>% 
  ggplot( aes(x = `% of Adults 25+ with College Degrees`, y = `Covid Deaths Per Thousand`, size = `Total Covid Deaths`, color=Quintile, text=County)) +
  geom_point(alpha = 6/10) +
  ggtitle("Higher Education Attainment in Michigan Counties by Covid Deaths Per Thousand" ) +
 theme(panel.background = element_rect(fill = "white",
                                colour = "beige",
                                size = 0.5, linetype = "solid")) +
         
  scale_color_manual(values = c("#999999", "#00798c", "#66a182", "#edae49", "#dc0000b2")) 




ggplotly(ed_sp )
```


```{r income plots}
income_sp <- covid_quintile %>% 
  ggplot( aes(x = `Median Household Income`, y = `Covid Deaths Per Thousand`, size = `Total Covid Deaths`, color=Quintile, text=County)) +
  geom_point(alpha = 6/10) +
  ggtitle("Median Household Income in Michigan Counties by Covid Deaths Per Thousand" ) +
 theme(panel.background = element_rect(fill = "white",
                                colour = "beige",
                                size = 0.5, linetype = "solid")) +
         
  scale_color_manual(values = c("#999999", "#00798c", "#66a182", "#edae49", "#dc0000b2")) 




ggplotly(income_sp )
```
