---
title: "Exploring the Data"
author: "brooke"
date: "7/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lets see what we can find out from the estimates (used for comparison between places)

I want to create five number summaries for comparison, isolate the variables I want for my table

https://www.statology.org/summary-table-in-r/


Here's how I made an aesthetically pleasing five number summary table of all numeric variables: 

1) Create a descriptive variable names, so I table looks more presentable
2) Pull a five number summary for all variables (grouped by county)
3) Round the data one decimal, for aesthetics
4) trim off unnecessary rows and columns
5) Write a CSV version of my table for future projects

Here's how I made it shiny! (link)


```{r load tidyverse}
library(tidyverse)
```

I want a version of my data with descriptive variable names rather than tidy ones -> I  am making two variable names so I can use this other table later
```{r}
table_covid_census <-  covid_census %>% 
  select( -Population) %>% 
  rename( 
    Population = "population19E",
    Households = "householdsE",
    `Median Age` = "median_ageE",
    `Median HHI` = "median_incomeE",
    `% A25+ College Grad+` = "bach_degree_plus_a25E",
    `% of HH on Public Assist.` = "public_assist_rate",
    `Unemployment Rate` = "unemployment_rate",
    `% Asain` = "percent_asian",
    `% Black` = "percent_black",    
    `% Native American` = "percent_native", 
    `% Pacific Islander` = "percent_pacific_islander",
    `% White` = "percent_white",            
    `% Hispanic` = "percent_hispanic",
    `Total Covid Deaths` = "Total_Deaths",
    `Covid Deaths Per Thousand` = "Deaths_Per_Pop_Thousand"
  )

table_covid_census
```



the describe function will give us a five number summary
pysch package conflicts with gg plot 


it will need to be installed 
install.packages("psych",dependencies=TRUE)

```{r five number sum it all}
sum_table_covid_census <- psych::describe(table_covid_census, fast = TRUE ) %>% 
  mutate_if(is.numeric, round,1)

sum_table_covid_census
```
#Change the first coluimn to the variable names, and remove the unneeded counts 83 for the counties of


```{r prepare data to be vizualized in a table}
sum_table_covid_census <- tibble::rownames_to_column(sum_table_covid_census, "VALUE") %>%
  select(-vars , -n)
 
```

update the data to a tibble so it's easier to move variables around
```{r}
 as_tibble(table_covid_census)
```


remove non-numerical values (you will get an error for the non-numerical values)
```{r remove the nonsense rows}
sum_table_covid_census <- sum_table_covid_census[-c(1, 2, 16, 17,18,19,20),]
sum_table_covid_census

```

put this in csv file
```{r}
write.csv(sum_table_covid_census,"~/STA 518/Stat-518-Portfolio/Final Project/Mi_Covid/data/covid_census_summary_stats.csv")
```
 
 the file can be used for a shiny tables page
 
 
I will clean up the table file now


```{r table_covid_census}

table_covid_census <-  table_covid_census %>% 
  select(County, Population, Households,`Total Covid Deaths`, `Covid Deaths Per Thousand`,  `Median Age`, `Median HHI`, `% A25+ College Grad+`,
         `Unemployment Rate`, `% of HH on Public Assist.`, `% Asain`, `% Black`, `% Native American`, `% Pacific Islander`, `% White`, `% Hispanic`)
```

 
 
```{r}
write.csv(table_covid_census,"~/STA 518/Stat-518-Portfolio/Final Project/Mi_Covid/data/Covid_Census_Table.csv")
```






