---
title: "Exploring the Data With Summary Stats"
author: "brooke"
date: "7/12/2022"
output: html_document
---


I created  five number summaries for all variables to provide a baseline of the state of Michigan, 
the data will be exported into a CVS file and transformed into an aesthetically pleasing dynamic table 

[Here's](https://073308-brooke.shinyapps.io/MIShiny/) the final table!


The script below:

1) Creates a descriptive variable names
2) Pulls a five number summary for all variables (grouped by county) using the ::pysch:: package
3) Rounds the data one decimal, for aesthetics
4) Trims off unnecessary rows and columns
5) Writes a CSV file for export



### Step 1
- install packages in console if needed, load the ::tidyverse:: 

```{r load tidyverse}
library(tidyverse)
```


### Step 2
- Rename variables with descriptive titles, allowing for a cleaner looking final table
- Select the numerical variables I can summarize

```{r rename variables}
table_covid_census <-  read.csv("~/STA 518/BrookemWalters-Portfolio/Stats 518 Final Project/data/covid_census.csv") %>% 
  select(County, population19E, householdsE, median_ageE,median_incomeE, bach_degree_plus_a25E, public_assist_rate, unemployment_rate, percent_asian, percent_asian, percent_black,percent_native, percent_pacific_islander, percent_white, percent_hispanic,Total_Deaths, Deaths_Per_Pop_Thousand) %>% 
  rename( 
    Population = "population19E",
    Households = "householdsE",
    `Median Age` = "median_ageE",
    `Median HHI` = "median_incomeE",
    `% A25+ College Grad+` = "bach_degree_plus_a25E",
    `% of HH on Public Assist.` = "public_assist_rate",
    `Unemployment Rate` = "unemployment_rate",
    `% Asian` = "percent_asian",
    `% Black` = "percent_black",    
    `% Native American` = "percent_native", 
    `% Pacific Islander` = "percent_pacific_islander",
    `% White` = "percent_white",            
    `% Hispanic` = "percent_hispanic",
    `Total Covid Deaths` = "Total_Deaths",
    `Covid Deaths Per Thousand` = "Deaths_Per_Pop_Thousand"
  )
glimpse(table_covid_census)
```

### Step 3

```{r remove county}
# I want county for another project, but I'll remove it here so I can summarize the numerical data
sum_table_covid_census <- select(table_covid_census, -County)
```


- Calculate a five number summary of all variable using the "describe" function from the ::pysch:: package
  - **note* the ::pysch:: has a major conflict the ::ggplot2:: -to avoid bugs later, it's best to run it as part of the function

```{r five number sum it all}
sum_table_covid_census <- psych::describe(sum_table_covid_census, fast = TRUE ) %>% 
  mutate_if(is.numeric, round,1)
glimpse(sum_table_covid_census)
```





### Step 4
- Change the first column to the variable names, for table readability 
- All of the variables are for all 83 counties, removing "n" column

```{r prepare data to be vizualized in a table}
sum_table_covid_census <- tibble::rownames_to_column(sum_table_covid_census, "socioeconomic factor")

sum_table_covid_census <-   select(sum_table_covid_census, -vars , -n)
```


### Step 5
- Write an CSV file for use import into the dynamic summary table

```{r write csv sum stat file}
write.csv(sum_table_covid_census,"~/STA 518/BrookemWalters-Portfolio/Stats 518 Final Project/Tables_Shiny/MIShiny/covid_census_summary_stats.csv")
```
 


### Step 7 
 - Clean up the full data set from Step 2 for use in the shiny table


```{r table_covid_census}
table_covid_census <-  table_covid_census %>% 
  select(County, Population, Households,`Total Covid Deaths`, `Covid Deaths Per Thousand`,  `Median Age`, `Median HHI`, `% A25+ College Grad+`,
         `Unemployment Rate`, `% of HH on Public Assist.`, `% Asian`, `% Black`, `% Native American`, `% Pacific Islander`, `% White`, `% Hispanic`)
```

 
### Step  8 
- Write a CSV file for use import into the dynamic county table 
```{r write the csv full table file}
write.csv(table_covid_census,"~/STA 518/BrookemWalters-Portfolio/Stats 518 Final Project/Tables_Shiny/MIShiny/Covid_Census_Table.csv")
```



##### [Part Three: Plotting the Data](https://rpubs.com/ekoorb03/Guides_Scatter_Plots)


###### ---[Back to Table of Contents](https://github.com/BrookemWalters/BrookemWalters-Portfolio#table-of-contents-brooke-walters-portfolio)---



