---
title: "Creating the dataset"
author: "brooke"
date: "7/12/2022"
output: html_document
---

# Create Retrieve, Wrangle, Mutate, Merge
Welcome to my project!


Steps taken to retrieve the most recent census and covid morbidity data, mutate variables into percentages, and then join the two datasets:


1) Retrieve the most up to date covid death totals for the state of Michigan by county
2) Calculate county covid deaths per 1000
3) Pull the most recent ACS demographic estimates (2019 is the most current release as of today)
4) Calculate, county rate of unemployment, public assistance usage, racial composition, etc
5) Merge the JH data with the ACS data





[Blrub about project]


### Step 0 : connect the census via *API* (insert proper usage here)
to connect to the census data request a key from: https://api.census.gov/data/key_signup.html
run the census_api_key code below
census_api_key("INSERT KEY HERE)", install = TRUE)
*you only have to do this one time*


### Step 1: install packages
#install.packages(c("tidycensus", "tidyverse", "plotly"))
#library(tidyverse)


```{r load tidyverse}
library(tidyverse)

```

## Part One

### Step 1.2:  pull the most recent data from John Hopkins github repostiory 
[ insert links and explanation of data]

```{r pull cummlative covid deaths from repository}

jh_covid_repo <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"
  
us_cd_deaths <- read_csv(jh_covid_repo)

```
### Step 1.3:  Filter the data by Michigan 

```{r isolate michigan counties from the data set}
mi_cd_deaths <- filter(us_cd_deaths, Province_State == "Michigan" )
```



### Step 1.4: 
```{r pull relevant variables into wide form}
# deaths are cumulative, pull the last column for the most current sum
mi_cd_deaths <- mi_cd_deaths %>% 
select(Admin2, Province_State, Lat, Long_, Population, FIPS, length(names(mi_cd_deaths)))

#update column names from dates to death, and change combined key to county for joins later
names(mi_cd_deaths)[7] <- "Total_Deaths"
names(mi_cd_deaths)[1] <- "County"
names(mi_cd_deaths)[2] <- "state_name"
names(mi_cd_deaths)[6] <- "county_fips"

```


### Step 1.5: 
```{r update total deaths to deathrate}
mi_cd_deaths <- mutate(mi_cd_deaths, Deaths_Per_Pop_Thousand =  (Total_Deaths / (Population / 1000 )))
mi_cd_deaths <- mutate_at(mi_cd_deaths, 8, round, 2)
#micdd_tidy is now ready to merge with census data

```

## Part Two

Step 2.1: Load Tidy Census
```{r}
library(tidycensus)
```


### Step 2.2: Pull data from ACS: # #https://data.census.gov/cedsci/table?g=0400000US26%240500000_0500000US26007&y=2020&d=ACS%205-Year%20Estimates%20Data%20Profiles&tid=ACSDP5Y2020.DP02
```{r get-data-from-the-2016-2020-5-year estimates ACS}
#find the following variables by county and create a csv for reference
v20<- load_variables(2020, "acs5", cache = TRUE)
write.csv(v20,"~/STA 518/Stat-518-Portfolio/Final Project/Mi_Covid/data/variable_names")

```

### Step 2.3: Create Function to select relevant estimates 
```{r estimate function }
estimates_only_by_GEOID <- function(x){
  x <- select(x, GEOID, ends_with("E")) 
}

```


"Health is influenced by many factors, which may generally be organized into five broad categories known as determinants of health: genetics, behavior, environmental and physical influences, medical care and social factors. These five categories are interconnected"

Pulling the cenusus data into two chunks so it's easier to follow

Chunk one ecconomic factors

### Step 2.4 Pull Factors 
population, and estimates: household, median age, median income, unemployed population, labor force population, households on public assistant,  and adults over 25 with a bachelor's degree or higher

```{r health determinate variables}
ac_eccon_factors <- get_acs(
  geography = "county",
  state = "MI",
  variables = c(population19 = "B01001_001",
                households = "B09001_002",
                median_age = "B01002_001",
                median_income = "B19013_001",
                unemployed_in_lf = "B23025_005",
                labor_force_pop = "B23025_002",
                hh_on_assistance = "B09010_002",
                bach_degree_plus_a25= "DP02_0068P"),
  output = "wide"
) 

```
### Step 2.5 Create Unemployment Rate, and public assistance rate



```{r pull social health determinants}
# use the function created above to select the relevant estimates 
ac_eccon_factors <- estimates_only_by_GEOID(ac_eccon_factors) %>% 
  #create the unemployment rate
  mutate(unemployment_rate = (unemployed_in_lfE / labor_force_popE)*100) %>% 
  mutate_at(11, round, 1) %>% 
  select(-labor_force_popE, -unemployed_in_lfE) %>% 

  # create the public assistance rate
  mutate(public_assist_rate = (hh_on_assistanceE / householdsE)*100) %>% 
  mutate_at(10, round, 1) %>% 
  select(-hh_on_assistanceE)

glimpse(ac_eccon_factors)

```

thoughts, I was unaware that nearly 1 third of Michigan is on public assistance!
quote stats

### Step 2.6 pull county racial composition

```{r Why do you have to bring race into it}
ac_race_factors <- get_acs(
  geography = "county",
  state = "MI",
  variables = c(population19 = "B01001_001",
                asian = "B03002_006",
                black = "B03002_004",
                native = "B03002_005",
                pacific_islander = "B03002_007",
                white = "B03002_003",
                hispanic = "B03002_012"),
  output = "wide"
)

```
```{r create racial percents}
#select estimates
ac_race_factors <- estimates_only_by_GEOID(ac_race_factors) %>% 
#create percentages  
  mutate(percent_asian = (asianE/ population19E)*100,
         percent_black = (blackE/ population19E)*100,
         percent_native = (nativeE/ population19E)*100,
         percent_pacific_islander = (pacific_islanderE/ population19E)*100,
         percent_white = (whiteE/ population19E)*100,
         percent_hispanic = (hispanicE/ population19E)*100) %>% 
  #round
  mutate_at(vars(starts_with("percent")), funs(round(., 1)))

#update variable to only reflect percentages
ac_race_factors <- ac_race_factors %>% 
  select(GEOID, (starts_with("percent",ignore.case = TRUE)))
glimpse(ac_race_factors)
```

this is where things get exciting!
time to merge the 2 census sets with the covid data by county!


```{r join both census sets}
ac_eccon_factors <-  left_join(ac_eccon_factors, ac_race_factors, by = "GEOID")
glimpse(ac_eccon_factors)
```
Change the "Name Field" (County, State) to just the county name, so the Census and Covid data can join 
```{r}
ac_eccon_factors <- ac_eccon_factors %>% 
  mutate(County = str_remove_all(NAME, " County, Michigan"))
glimpse(ac_eccon_factors)
```


The grand finale.... Merge everything together!

```{r join census and covid data}
covid_census <-  left_join(ac_eccon_factors, mi_cd_deaths, by = "County")
glimpse(covid_census)

# write csv as a backup
write.csv(covid_census,"~/STA 518/Stat-518-Portfolio/Final Project/Mi_Covid/data/covid_census.csv")
```

#Success!!! Data cleaning complete

