---
title: "Create Retrieve, Wrangle, Mutate, Merge"
author: "brooke"
date: "7/23/2022"
output: html_document
---
### Welcome to my project!




*The Script Below:* 

- Retrieves the most up-to-date covid death totals for the state of Michigan by county 
- Calculates county covid deaths per 1000 
- Pulls the most recent ACS demographic estimates (2019 is the most current release as of today) 
- Calculates, county rate of unemployment, public assistance usage, and racial compositions 
- Merges the [JHU CSSE COVID-19](https://github.com/CSSEGISandData/COVID-19) morbidity data with the [American Community Survey (ACS)](https://www.census.gov/data/developers/data-sets/acs-5year.html) estimates into a CSV file for export

[Here](https://github.com/BrookemWalters/BrookemWalters-Portfolio/blob/main/Stats%20518%20Final%20Project/Introduction.md#introduction-to-michigan-counties-in-a-pandemic) explains the inspiration and the potential bias I bring to my data analysis. 


### Step 0 : connect the census via **API**
[(full API directions here)](https://www.census.gov/programs-surveys/acs/data/data-via-api.html)

- To connect to the census data request a key from:[api.census.gov](https://api.census.gov/data/key_signup.html)
- Run the census_api_key code below: 
  - census_api_key("INSERT KEY HERE)", install = TRUE)
  
  (*you only have to do this one time*)

## **Part One** Retrieving the Lastest Covid Deaths by County,


### Step 1: 
- install packages in console if needed, load the ::tidyverse::

#install.packages(c("tidycensus", "tidyverse"))


```{r load tidyverse}
library(tidyverse)

```



### Step 1.2:  
 - pull the most recent data from [JHU CSSE COVID-19 Github repositiory](https://github.com/CSSEGISandData/COVID-19.)
 - and assign it a value

```{r pull cummlative covid deaths from repository}

jh_covid_repo <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"
  
us_cd_deaths <- read_csv(jh_covid_repo)

```


### Step 1.3:  
- filter the data by Michigan, and create a new value 

```{r isolate michigan counties from the data set}
mi_cd_deaths <- filter(us_cd_deaths, Province_State == "Michigan" )
```


### Step 1.4: 
 - trim the unnecessary variables
 - update variable names for common join fields, and future plots 
 
```{r pull relevant variables into wide form}
# deaths are cumulative, pull the last column for the most current sum
mi_cd_deaths <- mi_cd_deaths %>% 
select(Admin2, Province_State, Lat, Long_, Population, FIPS, length(names(mi_cd_deaths))) %>% 

#update column names from dates to death, and change combined key to county for joins later
  rename(  
  County = "Admin2",
  state_name = "Province_State",
  county_fips = "FIPS"
  )

names(mi_cd_deaths)[7] <- "Total_Deaths" 
```


### Step 1.5: 
 - create a covid death rate per thousand (mortality rate), around round to the nearest hundredths place

```{r update total deaths to deathrate}
mi_cd_deaths <- mutate(mi_cd_deaths, Deaths_Per_Pop_Thousand =  (Total_Deaths / (Population / 1000 )))
mi_cd_deaths <- mutate_at(mi_cd_deaths, 8, round, 2)
#micdd_tidy is now ready to merge with census data

```



## Part Two
### Step 2.1: Load Tidy Census

```{r load tidy census}
library(tidycensus)
```


### Step 2.2: 
- pull data from [ACS:](https://data.census.gov/cedsci/table?g=0400000US26%240500000_0500000US26007&y=2020&d=ACS%205-Year%20Estimates%20Data%20Profiles&tid=ACSDP5Y2020.DP02)
- assign data value "v20"
- capture all of the variable names in CSV format for reference

```{r get-data-from-the-2016-2020-5-year estimates ACS}
#find the following variables by county and create a csv for reference
v20<- load_variables(2020, "acs5", cache = TRUE)
write.csv(v20,"~/STA 518/Stat-518-Portfolio/Final Project/Mi_Covid/data/variable_names")

```

### Step 2.3: 
- create function to select relevant estimates to automate 1 step of trimming out unused variables

```{r estimate function }
estimates_only_by_GEOID <- function(x){
  x <- select(x, GEOID, ends_with("E")) 
}

```


For steps 2.4 - 2.6 I'll pull the ACS variables in two chunks, it's easier to manage the calculations this way
- set one economic factors
- set two racial population estimates

### Step 2.4 
 - pull factors:   population, and estimates: household, median age, median income, unemployed population, labor force population, households on public assistant,  and adults over 25 with a bachelor's degree or higher

```{r health determinate variables}
ac_eccon_factors <- get_acs(
  geography = "county",
  state = "MI",
  variables = c(population19 = "B01001_001",
                households = "B09001_002",
                median_age = "B01002_001",
                median_income = "B19013_001",
                unemployed_in_lf = "B23025_005",
                labor_force_pop = "B23025_002",
                hh_on_assistance = "B09010_002",
                bach_degree_plus_a25= "DP02_0068P"),
  output = "wide"
) 

```


### Step 2.5 
- create unemployment and public assistance rates, 
- remove the counts
- round for aesthetics 

```{r pull social health determinants}
# use the function created above to select the relevant estimates 
ac_eccon_factors <- estimates_only_by_GEOID(ac_eccon_factors) %>% 
  #create the unemployment rate
  mutate(unemployment_rate = (unemployed_in_lfE / labor_force_popE)*100) %>% 
  mutate_at(11, round, 1) %>% 
  select(-labor_force_popE, -unemployed_in_lfE) %>% 

  # create the public assistance rate
  mutate(public_assist_rate = (hh_on_assistanceE / householdsE)*100) %>% 
  mutate_at(10, round, 1) %>% 
  select(-hh_on_assistanceE)

glimpse(ac_eccon_factors)

```


### Step 2.6 
- pull county racial population variables

```{r Why do you have to bring race into it}
ac_race_factors <- get_acs(
  geography = "county",
  state = "MI",
  variables = c(population19 = "B01001_001",
                asian = "B03002_006",
                black = "B03002_004",
                native = "B03002_005",
                pacific_islander = "B03002_007",
                white = "B03002_003",
                hispanic = "B03002_012"),
  output = "wide"
)

```


### 2.7
 - create population rates by racial composition
 - round the values for consistency 
 - **Note *GEOID* will be used to join the two sets of ACS variables**
```{r create racial percents}
#select estimates
ac_race_factors <- estimates_only_by_GEOID(ac_race_factors) %>% 
#create percentages  
  mutate(percent_asian = (asianE/ population19E)*100,
         percent_black = (blackE/ population19E)*100,
         percent_native = (nativeE/ population19E)*100,
         percent_pacific_islander = (pacific_islanderE/ population19E)*100,
         percent_white = (whiteE/ population19E)*100,
         percent_hispanic = (hispanicE/ population19E)*100) %>% 
  #round
  mutate_at(vars(starts_with("percent")), funs(round(., 1)))

#update variable to only reflect percentages
ac_race_factors <- ac_race_factors %>% 
  select(GEOID, (starts_with("percent",ignore.case = TRUE)))
glimpse(ac_race_factors)
```


this is where things get exciting!
time to merge the 2 census sets with the covid data by county!

## Step 2.8
- join the eccomonic factors with the racial estimates

```{r join both census sets}
ac_eccon_factors <-  left_join(ac_eccon_factors, ac_race_factors, by = "GEOID")
glimpse(ac_eccon_factors)
```


## Step 2.9
- split the "Name Field" (County, State) to remove the state, allowing the ACS and the JHU CSSE COVID-19 data to join with a common variable and field name
```{r}
ac_eccon_factors <- ac_eccon_factors %>% 
  mutate(County = str_remove_all(NAME, " County, Michigan"))
glimpse(ac_eccon_factors)
```


The grand finale.... Merge everything together!

```{r join census and covid data}
covid_census <-  left_join(ac_eccon_factors, mi_cd_deaths, by = "County")
glimpse(covid_census)

# write csv as a backup -> update file path if needed
write.csv(covid_census,"~/STA 518/BrookemWalters-Portfolio/Stats 518 Final Project/data/covid_census.csv")
```

### Success!!! Data cleaning complete!

##### [Part Two: Summarizing the Data](http://rpubs.com/ekoorb03/Guides_Summary_Stats)


###### ---[Back to Table of Contents](https://github.com/BrookemWalters/BrookemWalters-Portfolio#table-of-contents-brooke-walters-portfolio)---

